\documentclass[11pt]{article}
\usepackage[top = 1in, bottom = 1in, left =1in, right = 1in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{tabu}
\usepackage{amssymb}
\usepackage{etoolbox}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{tikz-cd}
\usepackage{tikz}
\AtBeginEnvironment{proof}{\color{blue}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{fancyhdr}\pagestyle{fancy}
\newcommand{\la}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\tors}{\mathrm{tors}}
\newcommand{\ab}{\mathrm{ab}}
\newcommand{\Aut}{\operatorname{Aut}}
\newcommand{\Inn}{\operatorname{Inn}}
\newcommand{\im}{\operatorname{im}}
\newcommand{\lcm}{\operatorname{lcm}}
\newcommand{\ch}{\operatorname{char}}

%Math blackboard:
\newcommand{\bC}{\mathbb{C}}
\newcommand{\bF}{\mathbb{F}}
\newcommand{\bN}{\mathbb{N}}
\newcommand{\bQ}{\mathbb{Q}}
\newcommand{\bR}{\mathbb{R}}
\newcommand{\bS}{\mathbb{S}}
\newcommand{\bZ}{\mathbb{Z}}

%Greek blackboard font:
\newcommand{\bmu}{\mbox{$\raisebox{-0.59ex}
  {$l$}\hspace{-0.18em}\mu\hspace{-0.88em}\raisebox{-0.98ex}{\scalebox{2}
  {$\color{white}.$}}\hspace{-0.416em}\raisebox{+0.88ex}
  {$\color{white}.$}\hspace{0.46em}$}{}}

\lhead{University of California, Berkeley}
\rhead{Math 113 Section 6, Spring 2020}

\begin{document}
\begin{center}
\Large {Homework 10}\\
\small {Due Friday, April 24th}
\end{center}
Recall the definition of the semidirect product.
\begin{definition}
  Let $H,K$ be groups, and $\varphi:K\to\Aut(H)$ a group homomorphism.  Denote the induced action of $K$ on $H$ by:
  \[k\cdot h = \varphi(k)(h).\]
  The \textit{semidirect product} of $H$ and $K$ with respect to $\varphi$ is the set $H\rtimes K = \{(h,k):h\in H,k\in K\}$, where multiplication is defined by the rule:
  \[(h_1,k_1)(h_2,k_2) = (h_1(k_1\cdot h_2),k_1k_2).\]
\end{definition}
\begin{enumerate}
  \item{
  Let's make sure that $H\rtimes K$ is a group.
  \begin{enumerate}
    \item Show that $(1,1)\in H\rtimes K$ is the identity.  (Remember you have to check both sides).
    \item Show that $(h,k)^{-1} = (k^{-1}\cdot h^{-1},k^{-1})$. (As above, you have to check both sides).
    \item Prove that multiplication is associative.
  \end{enumerate}
  }
  \item{
  Let $G_1,G_2,\cdots,G_n$ be groups.  Show that:
  \[Z(G_1\times G_2\times\cdots\times G_n) = Z(G_1)\times Z(G_2)\times\cdot\times Z(G_n).\]
  Conclude that a product of groups is abelian if and only if the factors are.
  }
  \item{
  Let's classify some abelian groups!  List all \textit{abelian} groups of the following orders, in elementary divisor and invariant factor forms.
  \begin{enumerate}
    \item 100
    \item 243
    \item 9801
  \end{enumerate}
  }
  \item{
  Which of the following groups of order 80 are isomorphic?
  \begin{enumerate}
    \item $Z_5\times Z_4\times Z_4$
    \item $Z_{10}\times Z_8$
    \item $Z_4\times Z_{20}$
    \item $Z_8\times Z_5\times Z_2$
  \end{enumerate}
  }
  \item{
  Let $A$ be an abelian group of (invariant factor) type $(n_1,n_2,\cdots,n_s)$.  Show that there exists some element in $A$ of order $m$ if and only if $m|n_1$.  Conclude that the exponenet of $A$ is $n_1$.
  }
\end{enumerate}
\end{document}
